---
title: Creating complex dosing regimens with event objects
author: ~
date: '2017-08-27'
slug: creating-complex-dosing-regimens-with-event-objects
categories: []
tags: []
output: 
  blogdown::html_page:
    toc: true
    number_sections: true
---


```{r,echo=FALSE}
knitr::opts_chunk$set(comment='.', warning=FALSE, message=FALSE)
```


# Introduction
With this blog post, we'd like to introduce some new functions
in `mrgsolve` that work with event objects. The focus 
for this functionality is creating complicated dosing
regimens from simpler, elementary event objects.  And 
in particular, we'll see in this blog post how to 
easily create regular/irregular dosing patterns.
 
Let's get started

```{r,message=FALSE}
library(dplyr)
library(mrgsolve)
mod <- mread("pk2cmt", modlib()) %>% update(delta = 0.1) %>% Req(CP)
```

# Event objects
Recall that event objects are our way of creating simple
dosing interventions.  They are created with the `ev` function
and they look like a single dosing row from a data set.  

For example, we can have 100 mg daily dosing over a week like this
```{r}
e1 <- ev(amt = 100, ii = 24, addl = 6)
e1
```

```{r}
mod %>% mrgsim(events = e1, end = 2*168) %>% plot
```


# Sequences of event objects

Now, let's imagine a alternating dosing sequence of 100 mg daily for a 
week, then 200 mg daily for a week etc.  We can schedule this 
with a sequence of event objects.

First, make the 200 mg event object
```{r}
e2 <- ev(amt = 200, ii = 24, addl = 6)
```

Then, schedule them in a sequence

```{r}
e3 <- seq(e1, e2, e1)
e3
```

When we create a sequence of event objects, `mrgsolve` calculates
when each event is finished and automatically times each event
according to the end time from the preceeding event.  Notice
that each event starts one dosing interval after the last
dose of the previous event.

And try that out
```{r}
mod %>% mrgsim(events = e3, end = 5*168) %>% plot
```

## Using `wait` in event sequence

What if we want a 3-day holiday between each treatment period?  Simply
schedule a waiting period in the sequence:

```{r}
e4 <- seq(e1, wait = 72, e2, wait = 72, e1)
e4
```
Notice that the event objects (`e1` and `e2`) are 
unnamed arguments.  But `wait` is a named argument
with a single numeric value.  Introducing a `wait` period
in the sequence pushes the start of the next period 
in the sequence forward in time.

```{r}
mod %>% mrgsim(events = e4, end = 6*168) %>% plot
```

To code a cycle of 3 weeks on, one week off for two months

```{r}
e5 <- ev(amt = 200, ii = 24, addl = 20)

e6 <- seq(e5, wait = 168, e5)

mod %>% mrgsim(events = e6, end = 72*24) %>% plot
```


# Repeating a complicated regimen
Coding the 3 weeks on / 1 week off regimen for 2 months 
wasn't too bad, but it might get tedious if we need to do it for 
a year.

To make that simpler we can repeat the basic regimen 
as many times as we'd like with `ev_rep`

```{r}
e7 <- ev_rep(e5, n = 12, wait = 168)

mod %>% mrgsim(data = e7, end = 52*168) %>% plot(CP~time/7/24)
```

In `ev_rep`, there is a single `wait`ing period in between replicates
of the 3 week on period.

Alternatively, we could work with the 100/200/100 sequence
like this
```{r}
e4
e8 <- ev_rep(e4, n = 10, wait = 168)
mod %>% mrgsim(data = e8, end = 52*168) %>% plot(CP~time/7/24)
```


# Expand an event into a population
So far we have been dealing with a single ID in the event object
or data set.  The `ev_rep` function can easily expand that event
in to several IDs


```{r}
e9 <- seq(e1, wait = 72, e2)
e9 <- ev_rep(e9, id = 1:5)
e9
```

Now we have 5 patients each getting 100 mg daily for one week
and then 200 mg daily for a week after a 3 day break. 

```{r}
mod <- mread("popex", modlib()) %>% update(delta=0.1)
mod %>% mrgsim(data = e9, end = 600) %>% plot(DV~.)
```
# Environment
```{r,echo=FALSE}
devtools::session_info()
```

