---
title: Modeled interventions in mrgsolve
author: ~
date: '2018-09-02'
slug: modeled-interventions-in-mrgsolve
tags: ["events", "in-development"]
---

# Introduction

This post is to introduce modeled interventions in mrgsolve. The main use
case is to force mrgsolve to advance the 
system to a specific time so that some aspect of the system 
can change at that time.  This is similar to the `MTIME` functionality
that NONMEM provides.  Doses (bolus or infusions)
can also be implemented as a modeled event.  These events are "modeled"
because they originate from code in the `$MAIN` block, not the 
input data set, so that the events can be reactive to model parameters
or the state of the model itself. 

# Status
At the time of this post, this functionality is only in a development
branch on the mrgsolve GitHub.  But we've been working on and 
revising this concept for a while now, so hopefully it is 
more or less stable with a reasonable interface. That said, specifics
or syntax may change before this gets rolled out to CRAN. 

# Example: time-varying KA

One common use case for modeled events are a time-varying 
model parameter such as an absorption rate constant.  The model with 
__constant__ `KA` looks like this

```{r}
code <- '
$PARAM CL = 1, V = 20, KA1 = 0.1, KA2 = 0.5, KA3 = 5

$PKMODEL cmt = "GUT CENT", depot = TRUE

$PREAMBLE capture KA = KA1;

$MAIN
KA = KA2;
' 
```


```{r}
library(mrgsolve)

mcode("mevent0", code) %>% 
  ev(amt = 100) %>% 
  mrgsim(end = 24, delta = 0.1) %>%
  plot(CENT+KA~time)
```

For the model with __time-varying__ parameter, we want `KA` to start with a lower
value at the time of the oral
dose and then increase at some time after the dose.  And we 
want that to happen for every dose.  We model the time after the 
dose when we want KA to change and force the simulation to stop at that
time so we can make the required changes.

The model with time-varying `KA` might look like this

```{r}
code <- '
$PARAM CL = 1, V = 20, KA1 = 0.1, KA2 = 0.5, KA3 = 5

$PKMODEL cmt = "GUT CENT", depot = TRUE

$PREAMBLE capture KA = KA1;

$MAIN

if(EVID==1) {
  KA = KA1;
  self.mevent(TIME + 2, 33);
}

if(EVID==33) KA = KA3;

' 
```

The arguments for `self.mevent()` in this case are:

1. time
1. evid


Whenever we see a dose (`EVID==1`), we set `KA` to the initial (slower)
`KA` value and register a model event that in this case happens 
two hours later.  We use `EVID=33` for that change point event.  This is a 
arbitrary choice for `EVID` (it could be [almost] anything).  The idea here 
is that, when we see `EVID==33` come along, we change the `KA` to the faster value. 

We don't have plans to offer indicators like `MPAST` or `MTDIFF` as 
NONMEM does.  But we can achieve similar behavior by tagging the event
and then looking for that tag again down the road.  

The simulation looks like this

```{r}
mod <- mcode("mevent", code)

mod %>% 
  ev(amt = 100) %>% 
  mrgsim(end = 24, delta = 0.1) %>%
  plot(CENT+KA~time)
```

We can refine this even more with an additional intervention time for an 
intermediate value so that `KA` changes twice as the dose is absorbed

```{r}
code <- '
$PARAM CL = 1, V = 20, KA1 = 0.1, KA2 = 0.5, KA3 = 5

$PKMODEL cmt = "GUT CENT", depot = TRUE

$PREAMBLE capture KA = KA1;

$MAIN

if(EVID==1) {
  KA = KA1;
  self.mevent(TIME + 1, 33);
  self.mevent(TIME + 2, 34);
}

if(EVID==33) KA = KA2;
if(EVID==34) KA = KA3;

' 
```


We use a separate `EVID` (34) here to distinguish the second change 
from the first change (with `EVID 33`).


```{r}
mod <- mcode("mevent2", code)

mod %>% 
  ev(amt = 100) %>% 
  mrgsim(end = 8, delta = 0.1) %>%
  plot(CENT+KA~time)
```

Now we have the two change points implemented. 

Multiple doses would look like this

```{r}
mod %>% 
  ev(amt = 100, ii = 24, addl = 2) %>% 
  mrgsim(end = 120, delta = 0.1) %>%
  plot(CENT+KA~time)
```

And we can model this so that the change point is a function of a model 
parameter

```{r}
code <- '
$PARAM CL = 1, V = 20, KA1 = 0.1, KA2 = 0.5, KA3 = 5
MTIME1 = 2


$PKMODEL cmt = "GUT CENT", depot = TRUE

$PREAMBLE capture KA = KA1;

$MAIN

if(EVID==1) {
  KA = KA1;
  self.mevent(TIME + MTIME1, 33);
}

if(EVID==33) KA = KA3;

' 
```

```{r}
mod <- mcode("mevent3", code)

idata <- data.frame(MTIME1 = seq(0,2.5,0.5))

mod %>% 
  ev(amt = 100) %>% 
  idata_set(idata) %>%
  mrgsim(end = 8, delta = 0.1) %>%
  plot(CENT+KA~time)
```



# Example: modeled dosing events

In the previous example, we used the `mevent` call to 
to force the simulation to stop at the time we wanted a model 
parameter to change. This was accomplished by an event object
with `EVID=2`, which forces the solver to stop, reset, and then keep going 
with the new model parameters.

We can also use this event mechanism to implement actual doses.  Now
at the change point, we add some mass to one of the compartments.  An example
model for q12 hour dosing might look like:

```{r}
code <- '
$PARAM CL = 1, V = 20, KA = 2

$PKMODEL cmt = "GUT CENT", depot = TRUE

$MAIN

if(TIME/12 ==floor(TIME/12) && TIME < 120) {
  self.mevent(TIME, 1, 1, 100, 0);
}

' 
```


The arguments for `self.mevent()` in this case are:

1. time
1. evid
1. compartment
1. dose amount
1. rate (for infusions)


```{r}
mod <- mcode("mevent4", code)

mod %>%  mrgsim(end = 240, delta = 0.2) %>% plot(CENT~time)
```

So here, we simulated without the data set and implemented the 
dosing from within the model.  This isn't the way you should be implementing
dosing like this on a regular basis, but it shows how you can 
trigger doses from within the model based on what has gone on up to that 
point in the simulation. 





