---
title: Model Gallery
date: 2017-02-23T08:00:00
type: page
output: 
  blogdown::html_page: 
    toc: TRUE
    toc_depth: 1
tags: "events"
---


```{r,echo=FALSE}
knitr::opts_chunk$set(comment='.', fig.align="center",message=FALSE,cache=TRUE, autodep=TRUE)
```


```{r, message=FALSE}
library(lattice)
library(ggplot2)
library(dplyr)
```


# Load packages required for the gallery
```{r,message=FALSE}
library(mrgsolve)
library(dplyr)
library(ggplot2)
library(magrittr)
library(tidyr)
```



# Simple PK model

About this model specification code

  * We set a simulation grid of 0.1 units in `$SET`
  * Compartment number (2) and names (`GUT` and `CENT`) are established in `$CMT`
  * For dosing purposes, `GUT` is compartment 1 and `CENT` is compartment 2 due to the order of their declaration
  * Compartment names and numbers can be checked with `init(mod)`

```{r}
code <- '
$PARAM CL = 1.3, VC=28, KA=0.6

$SET delta=0.1

$CMT GUT CENT

$ODE
dxdt_GUT = -KA*GUT;
dxdt_CENT = KA*GUT - (CL/VC)*CENT;

$TABLE double CP = CENT/VC;

$CAPTURE CP
'
```



Compile and load
```{r}
mod <- mread("gallery1", tempdir(),code)


mod
```


## Run with initial value
```{r}
out <- mod %>% init(GUT=1000) %>% mrgsim(end=72)

plot(out)

out
```


## Run with `events` object

```{r}
out <- mod %>% ev(amt=1000, ii=12, addl=15) %>% mrgsim(end=336)

plot(out)
```

The events object was just a simple data set that looks like this
```{r}
dosing <- events(out)
dosing
```


## Run with data set

Just like the events object, but we can make it more complicated, specify more people, etc...

```{r}
data <- expand.ev(amt=c(300,1000,3000), ii=12, addl=6) %>% mutate(dose=amt)
data
```


```{r}
out <- mod %>% data_set(data) %>% carry.out(dose) %>% Req(CP) %>% mrgsim(end=168)
```

```{r}
plot(out, CP~time|factor(dose),scales="same")
```


# PK model with different bioavability and dosing lag time


About this model specification code

  * Set `F_CENT` and `ALAG_CENT` in `$MAIN`
  * Since `F_CENT` is set to `F1` and `ALAG_CENT` is set to `ALG`; since `F1` and `ALG` are named in `$PARAM`, we can update
  them on the `R` side either through `param` or in an input data set that is also determined on the `R` side of things


```{r}
code <- '
$PARAM CL = 1.3, VC=28, KA=0.6, ALG=0, F1=1,FORM=1

$SET delta=0.1

$CMT CENT

$MAIN
F_CENT = 1;

// Formulation 2 has lower bioavailability
if(FORM==2) F_CENT = F1;

// Same lag time across formulations
ALAG_CENT = ALG;


$ODE dxdt_CENT = - (CL/VC)*CENT;

$TABLE double CP = CENT/VC;

$CAPTURE CP

'
```



Compile and load
```{r}
mod <- mread("gallery2", tempdir(), code)
```


Dose into the central compartment.  Also, `ALG`, `F1`, and `FORM` are updated via
the data set and therefore need to be columns in the data set and listed in `$PARAM`

```{r}
data <- 
  expand.ev(amt=1000, ALG=seq(0,12,4),FORM=c(1,2)) %>% 
  mutate(F1 = ifelse(FORM==1,1,0.5))

data

out <- mod %>% data_set(data) %>% carry.out(FORM,ALG) %>% mrgsim(end=72)

plot(out, CP~time|factor(FORM),group=ALG, 
     scales="same", auto.key=list(columns=4))
```


# PK model with residual error

About this model specification code 

  * `$SIGMA` sets the variance of `EPS(n)`
  * Setting `end` to a value less than 0 will signal `mrgsolve` to skip the simulation time grid
  * We specify `add` to get a specific set of simulation times (not necessarily evenly spaced)


```{r}
code <- '
$PARAM CL = 1.3, VC=28, KA=0.6

$GLOBAL #define CP (CENT/VC)

$SIGMA 0.025

$SET end=-1, add=c(0,0.25,0.5,1,2,4,6,8,10,12,16,24,36,48)

$CMT GUT CENT

$ODE
dxdt_GUT = -KA*GUT;
dxdt_CENT = KA*GUT - (CL/VC)*CENT;

$CAPTURE CP DV

$TABLE double DV = CP*exp(EPS(1));
'

mod <- mread("ruv", tempdir(), code) 
```


```{r}
set.seed(3203)
out <- 
  mod %>% 
  init(GUT=1000) %>% 
  obsonly %>% 
  mrgsim(Req="CP,DV") %>% as.data.frame

ggplot(data=out, aes(x=time)) + 
  geom_point(aes(y=DV),size=3) + 
  geom_line(aes(y=CP), col="firebrick",lwd=1)

```



# Indirect response model driven by continuous PK

About this model specification code

  * `CP` is "defined" in `$GLOBAL`
  * `INH` is declared as a double precision variable in `$ODE`
  * `INH` will be available globally (`mrgsolve` will declare the variable in `$GLOBAL` for you)
  * `INH` can be changed at any time, but `CP` cannot
  * Both `CP` and `INH` can be inserted into output in `$TABLE` or `$CAPTURE`


```{r}
code <- '
$PARAM CL = 5.3, VC=28, KA=0.6, KIN=20, KOUT=0.01, IC50=10

$SET delta=0.1

$CMT GUT CENT RESPONSE

$GLOBAL
#define CP (CENT/VC)

$MAIN
RESPONSE_0 = KIN/KOUT;

$ODE

double INH = CP/(IC50+CP);

dxdt_GUT = -KA*GUT;
dxdt_CENT = KA*GUT - (CL/VC)*CENT;

dxdt_RESPONSE = KIN*(1-INH) - KOUT*RESPONSE;


$CAPTURE CP INH
'

mod <- mread("gallery3", tempdir(), code)

out <- mod %>% 
  ev(time=120,amt=1500,rate=5) %>% 
  Req(CP,RESPONSE) %>% mrgsim(end=480*2)

plot(out)
```


# Indirect response model driven by metric summarizing exposure

About this model specification code

  * Take out the PK model 
  * `AUC` will drive the PD model
  * `AUC` is declared in `$PARAM`; we will manipulate that value (AUC>0 for on treatment, AUC==0 for off treatment) through 
  the input data set

```{r}

code <- '
$PARAM AUC=0, KIN=100, KOUT=0.2, AUC50=10

$SET delta=0.1

$CMT RESPONSE

$GLOBAL
#define INH (AUC/(AUC50+AUC))

$MAIN
RESPONSE_0 = KIN/KOUT;

$ODE

dxdt_RESPONSE = KIN*(1-INH) - KOUT*RESPONSE;

$CAPTURE AUC
'

mod <- mread("gallery5", tempdir(), code)

```


Make a template data set with `evid`=2 and `AUC` to drive the PD model

```{r}
data <- 
  ev(ID=1,evid=2,AUC=15, ii=48,addl=0,amt=0) %then% 
  ev(ID=1,evid=2,AUC=0,amt=0)

data

out <- mod %>% data_set(data) %>% obsonly %>% mrgsim(end=120)

plot(out)
```


# Effect compartment model


About this model specification code

  * This code replicates  model PD21 in Gabrielsson and Weiner 2007.



```{r}
code <- '
$PARAM  
VC = 2.44, KA = 0.92, K10 = 0.44, K12 = 0.36, K21 = 0.24
E0 = 157, EMAX = 29.95, EC50=4.956, KEO=1.305

$SET delta=0.1, end=36

$CMT GUT CENT PERIPH CE

$ODE 
double CP = CENT/VC;

dxdt_GUT = -KA*GUT;
dxdt_CENT = KA*GUT - K10*CENT - K12*CENT + K21*PERIPH;
dxdt_PERIPH = K12*CENT - K21*PERIPH;
dxdt_CE = KEO*(CP-CE);

$TABLE
capture EFFECT = E0 - EMAX*CE/(CE+EC50);
CP = CENT/VC;

$CAPTURE EFFECT CP
'

mod <- mread("effcmt", tempdir(), code)

```


Simulate a snigle 200 unit dose

```{r}
out <- mod %>% ev(amt=200) %>% obsonly %>% mrgsim(Req="CP,CE,EFFECT")

```

```{r,fig.height=4}
plot(out)
```


Simulate steady state PK/PD under Q12 hour dosing

```{r}
out <- 
  mod %>% 
  ev(amt=200,ss=1,ii=12,addl=10)%>%
  mrgsim(Req="CP,CE,EFFECT", end=168)

plot(out)
```




# Population PK model with covariates and `IOV` 

About this control stream

  * `IOV` is declared as double precision variable that is set to either `ETA(4)` or `ETA(5)`
  * We need to make sure that $\omega^2_{4}$ and $\omega^2_{5}$ are equal
  * We assign different names to the first and second `$OMEGA` matrices so we can update them by name later
  * We are tabling out some `ETAs` to look at them later as well; this is for demonstration purposes

```{r}
code <- '
$PARAM TVCL = 1.3, TVVC=28, TVKA=0.6, WT=70, OCC=1

$SET delta=0.1

$CMT GUT CENT

$MAIN

double IOV = EOC1;
if(OCC==2) IOV = EOC2;

double CLi = exp(log(TVCL) + 0.75*log(WT/70) + ECL + IOV);
double VCi = exp(log(TVVC) + EVC);
double KAi = exp(log(TVKA) + EKA);

$OMEGA 
@name IIV @labels ECL EVC EKA
0 0 0
$OMEGA @name IOV @labels EOC1 EOC2
0 0

$SIGMA 0

$ODE
dxdt_GUT = -KAi*GUT;
dxdt_CENT = KAi*GUT - (CLi/VCi)*CENT;

$TABLE
double CP = CENT/VCi;
double ETA1 = ETA(1);
double ETA2 = ETA(2);

$CAPTURE IOV CP ETA1 ETA2
'

mod <- mread("gallery4", tempdir(), code)

```



Set up dosing events; occasion 2 is the second dose

```{r}
dose <- ev(amt=300,OCC=1) + ev(amt=300,time=48,OCC=2)
dose
```



ETAs are all zeros per the model specification code
```{r}
out <- mod %>% ev(dose) %>% mrgsim(end=120, nid=5)

plot(out)
```



Update `$OMEGA`

```{r}
mod <- mod %>% omat(IIV=cmat(0.1, 0.45, 0.3, 0,0,0), IOV=dmat(0.5,0.5))
```



And check the result
```{r}
revar(mod)
```

Simulate and plot
```{r}
out <- mod %>% ev(dose) %>% mrgsim(end=120,nid=5,seed=9765)
```


ETAs
```{r}
plot(out,ETA1+ETA2+IOV~time|factor(ID),scales="same", layout=c(5,3))
```

PK
```{r}
plot(out,CP~time|factor(ID),scales="same", layout=c(5,1))
```



Just look at WT with no IIV/IOV
```{r}
mod <- mod %>% drop.re
```

```{r}
data <- expand.ev(amt=300,ii=12, addl=19,WT=seq(40,140,20))

out <- mod %>% data_set(data) %>% Req(CP) %>% mrgsim(end=240, carry.out="WT")

out %>% 
  as.tbl %>%
  ggplot(data=., aes(x=time, y=CP,col=WT,group=ID)) + 
  geom_line(lwd=1)
```


Cmin versus weight
```{r}
out %>% subset(time==240) %>%
  ggplot(data=.,aes(x=WT,y=CP)) + 
  geom_line(lwd=1) + geom_point(size=3) + xlab("Cmin")
```




# PK with dialysis sessions 3 times per week
##'
About this control stream

  * `GUT` and `CENT` are regular compartment; we have to provide a differential equation for each
  * `DIAL` is a compartment that only holds a value; `dxdt_DIAL` will automatically be set to `0`
  * `CLadd` is the additional drug clearance while on dialysis



```{r}
code <- '
$PARAM CL=0.7, VC=16, KA=1.8, CLadd = 2.1

$CMT GUT CENT

$VCMT DIAL

$ODE
dxdt_GUT = -KA*GUT;

dxdt_CENT = KA*GUT -  (CL + CLadd*DIAL)*CENT/VC;

$TABLE double CP = CENT/VC;

$CAPTURE CP
'

```


```{r}
mod <- mread("dialysis", tempdir(), code) %>% update(end=336, delta=0.25)
```

__Data assembly for doses and dialysis sessions__

Set up three weeks of dosing
```{r}
doses <- expand.ev(amt=300, ii=24, addl=13)
```

MWF dialysis sessions
```{r}
mon <- expand.ev(time=0,ii=7*24,addl=2,cmt=3,evid=8,amt=1)
wed <- mon %>% mutate(time=2*24)
fri <- wed %>% mutate(time=4*24)
```


Dialysis starts 8 hours after the AM dose
```{r}
dial_start <- bind_rows(mon,wed,fri) %>% mutate(time=time+8)
```

Dialysis lasts 3 hours
```{r}
dial_end <- dial_start %>% mutate(time = time+3, amt=0)
```

Doses and dialysis
```{r}
events <- 
  bind_rows(dial_start,dial_end,doses) %>% 
  arrange(time) %>% as.data.frame %>% as.ev

events

```


__Simulate__
```{r}
out <- mod %>% ev(events) %>%  obsonly %>% mrgsim
```



```{r}
plot(out)
```



PK plot with dialysis sessions marked in red
```{r,echo=FALSE}
dial <- left_join(as.data.frame(events), as.data.frame(out)) %>% filter(cmt==3)
dial <- bind_rows(dial, filter(dial, DIAL==1) %>% mutate(CP=NA, time=time+0.1)) %>% arrange(time)

```

```{r}
out %>% as.tbl %>% filter(time <=7*24) %>%
  ggplot(data=., aes(x=time, y=CP)) + geom_line(lwd=1) + 
  geom_path(data=dial,col="firebrick", lwd=6, alpha=0.4)
```


# References
J. Gabrielsson and D. Weiner. Pharmacokinetic and Pharmacodynamic Data Analysis: Concepts and Applications, Fourth Edition. Taylor & Francis, 2007.




# Session Info
```{r}
devtools::session_info()
```


