---
title: Access Models in the Internal Library
date:  2017-02-23T10:13:36
tags: ["model library", "pd model"]
---

```{r, echo=FALSE}
library(knitr)
opts_chunk$set(comment='.',message=FALSE,fig.align="center",message=FALSE,cache=TRUE, autodep=TRUE)
```


```{r}
library(mrgsolve)
library(dplyr)
```


### PK model

Here, we name the name of the model we want from the library (`pk2cmt`) and 
we point `mread` to the `project` directory where the internal model library is 
stored.  This internal library is located wherever `mrgsolve` is installed.  The install 
location for `mrgsolve` is not obvious, so we provide a function (`modlib`) that 
returns the currect path to the model library.


```{r}
mod <- mread("pk2cmt", modlib())

mod %>%
  ev(amt=100,rate=3,addl=4,ii=48,cmt=2) %>%
  mrgsim(end=320) %>% 
  plot(CP~.)

```



```{r}
see(mod)
```




To get a list of available models
```{r}
?modlib
modlib(list=TRUE)
modlib()
```



### Viral model
```{r}
mod <- mread("viral1",modlib())

e <- 
  ev(amt=50, cmt="expos",time=2) + 
  ev(amt=0, cmt="expos", evid=8,time=11)

out <- 
  mod %>%
  ev(e) %>%
  update(end=28,delta=0.1) %>%
  knobs(delta=seq(0.2,0.8,0.1))


plot(out,logChange~time,groups=delta,auto.key=list(columns=4))
```

### PK/PD model
```{r}
mod <- mread("irm1", modlib())

see(mod)

mod %>% ev(amt=700,time=50) %>% param(n=1.5,KOUT=0.02) %>%
  mrgsim(end=480) %>% plot(CP+RESP~.)
```


### Population PK model
```{r}
mod <- mread("popex", modlib())
```


```{r}

mod %>%
  ev(amt=100) %>% 
  mrgsim(nid=100, end=96, delta=0.25) %>%
  plot(DV+ECL~.)
```
