---
date: 2017-02-23T16:49:01-06:00
title: Quick hit feature demos
output: 
  blogdown::html_page:
    toc: true
slug: quick
---

These are all miscellaneous features that may be difficult to find in other documentation.


```{r,echo=FALSE}
knitr::opts_chunk$set(message=FALSE,comment='.',cache=TRUE,autodeps=TRUE)
```

```{r,message=FALSE}
library(mrgsolve)
library(dplyr)
```


# `$` operator for model object gets the parameter value

If our model parameters are
```{r}
mod <- mrgsolve:::house()
param(mod)
```

we can pick a parameter value with
```{r}
mod$CL

mod$WT
```




# Zero all random effect variances on the fly

If your model has random effects, you can easily and temporarily zero them out. 
```{r}
mod <- mread("popex", modlib()) %>% update(end=96) %>% Req(DV,CL,V)

omat(mod)
```

It is easy to simulate either with or without the random effects in the simulation: this change can be made on the fly.

Use `zero_re` to make all random effect variances zero
```{r}
mod %>% zero_re %>% omat
```

By default, both `OMEGA` and `SIGMA` are zeroed.  Check the arguments for `zero_re` to see how to selectively zero `OMEGA` or `SIGMA`.

Compare the population output 

```{r}
mod %>% ev(amt=100) %>% mrgsim(nid=20) %>% plot
```

with 

```{r}
mod %>% ev(amt=100) %>% zero_re %>% mrgsim(nid=20) %>% plot
```




# Plot formulae
We commonly plot simulated output with a special plot method.  By default, you get all
compartments and output variables in the plot.

```{r}
mod <- mrgsolve:::house() %>% events(ev(amt=100))

mod %>% mrgsim %>% plot
```


The plot can be customized with a formula selecting variables to plot. Other arguments 
to `lattice::xyplot` can be passed as well.
```{r}
mod %>% mrgsim %>% plot(CP+RESP ~ time, lty=2, col="firebrick")
```

