---
output: html_document
title: "Flexible, heterogeneous simulation designs in a population"
date: 2017-01-05T13:09:02
---


Design lists help you assign different designs to 
different groups in a population or specific 
designs to specific individuals. 


```{r,echo=FALSE}
knitr::opts_chunk$set(comment='.',message=FALSE)
```


```{r}
library(mrgsolve)
library(dplyr)
```


### Assign designs to individuals

To illustrate, let's make a population of 
4 individuals, all with different simulation __end__ times.


```{r}
des <- data_frame(ID=1:4, end=seq(24,96,24))

des
```

For simplicity, we will only vary the simulation end time in 
this example.  See later examples where `start`, `delta` and 
`add` can varied as well. 


We can turn this in to a list of designs with `as_deslist`.  

```{r}
as_deslist(des, descol="ID")
```

`as_deslist` returns one design for each individual, one
for each unique level of `descol`. The 
deslist is a list of `tgrid` objects (see `?tgrid`).
Note also that `descol` is retained as an attribute to be used
later.

Let's set up a simulation that includes these 4 IDs; 
we load a model and, importantly, set up an `idata_set`
for the simulation that includes all 4 IDs in the design list.

```{r}

mod <- mrgsolve:::house() %>% ev(amt=100)

idata <- select(des,ID)
idata

des1 <- as_deslist(des,"ID")

```


When we run the simulation, pass in the 
design list to `design` in the pipeline


```{r}
out <- 
  mod %>%
  idata_set(idata) %>%
  design(des1) %>%
  mrgsim

```


We see that ID 1 has a 24 hour end time, ID 2 
has 48 hour simulation time, ID 3 with 72 hour 
simulation time, and ID 4 96 hour simulation time 
as reflected in the list of the designs.

```{r}
plot(out, CP~time|ID)
```



__Note__: Check the arguments to `design` (`?design`).  There
is a `descol` argument that is required.  `descol` in this function 
refers to a column in `idata_set` to be used as the grouping variable
to assign the sampling design.  `as_deslist` also had a `descol`
argument that referred to a column in the input data frame
for that function.  We don't need to pass `descol` to `design()`
because we created the design list with `as_deslist`: `design()`
reads `descol` from the attribute.  We don't __have__ to 
use `as_deslist` to create the design list.  It could be just a 
plan old `R` list created by you with `tgrid` objects.  In that 
case, you must state what `descol` is when you call `design()`.  
And it can't be emphasized enough here that you __MUST__ use
an `idata_set` for this to work and `idata_set` must 
contain a valid `descol`.  


### Assign designs to treatment arms or groups

Now, let's simulate a trial with 5 patients in each of
4 treatment arms.  In this trial,
arm 1 lasts 24 hours, arm 2 last 48 hours ... etc.  But
every patient with the arm 1 indicator will get simulated
for 24 hours, every patient with arm 2 indicator will get simulated
for 48 hours and so on.


```{r}
idata <- expand.idata(ARM=1:4,ID=1:5)

head(idata)

```

Now, let's setup the designs based on `ARM` rather 
than `ID`

```{r}

des <- distinct(idata,ARM) %>% mutate(end=seq(24,96,24))

des

```

The simulation works the same way

```{r}
set.seed(11)

out <- 
  mod %>% 
  idata_set(idata) %>%
  omat(dmat(1,1,1,1)/10) %>%
  design(as_deslist(des,"ARM")) %>% 
  mrgsim(carry.out="ARM")

plot(out, CP~time|factor(ARM))

plot(out, CP~time|factor(ID))

```

### `list-cols` and `additional` times

Hopefully it's clear that columns named
`start`, `end`, and `delta` in the 
the input data frame passed to `as_deslist`
are just numeric values that form 
the time grid object (see `?tgrid`).

What about `add`, the vector of ad-hoc times
for the simulation?  These, too, can be 
accommodated with a `list-col` column in 
the input data frame.  Note that `list-cols`
are only really supported in specialized
 `tibble` type data frames.



These are random times for IDs 1 and 2
```{r}
set.seed(12)
t1 <- c(0,sample(1:24,12))
t2 <- c(0,sample(1:96,12))

t1
t2
```

__Note__: When we simulate with `end < 0`, `mrgsolve` 
knows
to ignore `start/end/delta` and only use the 
times in `add` for observations.

```{r}
des <- data_frame(ID=1:2, end=-1, add=list(t1,t2))
des

mod %>% 
  idata_set(des, select=ID) %>%
  design(as_deslist(des)) %>% 
  mrgsim %>% 
  plot(CP~time|factor(ID), type='b')
```

Ok ... not the most lovely-looking result we've seen before, 
but maybe that's just what you needed in this simulation.

