---
title: "Work with event objects"
date: 2017-01-31T08:00:00
---


```{r setup, echo=FALSE,message=FALSE}
knitr::opts_chunk$set(echo = TRUE,comment='.',fig.align="center",message=FALSE,cache=TRUE, autodep=TRUE)
```

```{r setup2}
library(mrgsolve)
library(dplyr)
```


# Introduction

Event objects are simple ways to implement PK dosing events into your model simulation.  

# Setup
Let's illustrate event objects with a one-compartment, PK model.  We read this model from the `mrgsolve` internal model library.
```{r,message=FALSE}
mod <- mread("pk1cmt", modlib(), end=216, delta=0.1)
```

# Events

Events are constructed with the `ev` function

```{r}
e <- ev(amt=100, ii=24, addl=6)
```

This will implement 100 unit doses every 24 hours for a total of 7 doses.  `e` has class `ev`, but really it 
is just a data frame

```{r}
e
```

```{r}
as.data.frame(e)
```

We can implement this series of doses by passing `e` in as the `events` argument to `mrgsim`
```{r}
mod %>% mrgsim(events=e) %>% plot(EV1+CP~time)
```


The events can also be implemented with the `ev` constructor along the simulation pipeline
```{r}
mod %>%
  ev(amt=100, ii=24, addl=6) %>%
  mrgsim %>% 
  plot(CP~time)
```


# Event expectations

1. `amt` is required
1. `evid=0` is forbidden
1. Default `time` is 0
1. Default `evid` is 1
1. Default `cmt` is 1

Also by default, `rate`, `ss` and `ii` are 0.


# Combine events

`mrgsolve` has operators defined that allow you to combine events.  Let's first define some event objects.

```{r}
e1 <- ev(amt=500)
e2 <- ev(amt=250, ii=24, addl=4)
e3 <- ev(amt=500, ii=24, addl=0)
e4 <- ev(amt=250, ii=24, addl=4, time=24)
```


We can combine `e1` and `e3` with a collection operator

```{r}
c(e1,e4)
```

`mrgsolve` also defines a `%then$` operator that lets you execute one event and `%then%` a second event
```{r}
e3 %then% e2
```

Notice that `e3` has both `ii` and `addl` defined.  This is required for `mrgsolve` to know when to start `e2`.


# Combine event objects to create a data set

We can take several event objects and combine them into a single simulation data frame with the `as_data_set` function.

```{r}
e1 <- ev(amt=100, ii=24, addl=6, ID=1:5)
e2 <- ev(amt=50,  ii=12, addl=13, ID=1:3)
e3 <- ev(amt=200, ii=24, addl=2, ID=1:2)
```

When combined into a data set, we get 
* N=5 IDs receiving 100 mg Q24h x7 
* N=3 IDs receiving 50  mg Q12h x 14
* N=2 IDs receiving 200 mg Q48h x 3

```{r}
data <- as_data_set(e1,e2,e3)

data
```

To simulate from this data set, we use the `data_set` function.  First, let's load a population PK model

```{r,message=FALSE}
mod <- mread("popex", modlib())
```
```{r}
mod %>% data_set(data) %>% mrgsim(end=336) %>% plot(GUT+DV ~  .)
```















